---
title: "Report"
author: Anusha Srinivasan, Practicum II CS5200- Spring 2024
date: "17th April, 2024"
output:
  pdf_document: default
  html_notebook: default
---

```{r connection, message= FALSE, echo=FALSE, warning=FALSE}
library(RMySQL)
library(DBI)
library(ggplot2)
library(knitr)
library(dplyr)
library(tidyr)
# Database connection 
dbname <- 'dbname'  
host <- 'host'
username <- 'username'  
password <- 'password'  
port <- 
con <- dbConnect(RMySQL::MySQL(), dbname=dbname, host=host, username=username, password=password, port=port)
```

# Analytical Query I: Top five sales reps with the most sales broken down by year. 
```{r top-sales-reps-ranking, echo=FALSE, message=FALSE, comment=NA}
query1 <- "
SELECT year, rep_name, SUM(total_amount_sold) AS total_sales
FROM rep_facts
GROUP BY year, rep_name
ORDER BY year, total_sales DESC
"
top_reps_data <- dbGetQuery(con, query1)

# Use dplyr to rank reps within each year and select the top five
top_reps_ranked <- top_reps_data %>%
  group_by(year) %>%
  mutate(rank = dense_rank(desc(total_sales))) %>%
  filter(rank <= 5) %>%
  ungroup()

# Display the table using kable for each year
top_reps_by_year <- split(top_reps_ranked, top_reps_ranked$year)

tables <- lapply(names(top_reps_by_year), function(year) {
  year_data <- top_reps_by_year[[year]]
  kable(year_data, caption = paste("Top Five Sales Reps for", year), format = "markdown")
})

# Now print or return the tables
print(tables)
```
# Analytical Query II: Total sales per product per quarter. 
```{r total-sales-per-product, echo=FALSE, warning=FALSE}
product_sales_query <- "
SELECT year, product_name, quarter, SUM(total_amount_sold) AS total_sales
FROM product_facts
GROUP BY year, product_name, quarter
ORDER BY year, product_name, quarter
"
product_sales <- dbGetQuery(con, product_sales_query)
knitr::kable(product_sales, caption = 'Total Sales Per Product Per Quarter by Year')
product_sales$quarter <- factor(product_sales$quarter, levels = c("Q1", "Q2", "Q3", "Q4"))

# Generate the line plot 
ggplot(product_sales, aes(x = quarter, y = total_sales, group = product_name, color = product_name)) +
  geom_line() +
  geom_point() +
  facet_wrap(~year, scales = 'free_x') + 
  theme_minimal() +
  labs(title = 'Total Sales Per Product Per Quarter by Year',
       x = 'Quarter',
       y = 'Total Sales') +
  theme(legend.position = "bottom")  

```

# Analytical Query III: Number of units sold per product per region.
```{r units-per-product-region, echo=FALSE, warning=FALSE}
#query
query_units_per_product_region <- "
SELECT product_name, territory, SUM(total_units_sold) AS total_units
FROM product_facts
GROUP BY product_name, territory
ORDER BY product_name, territory
"

units_per_product_region <- dbGetQuery(con, query_units_per_product_region)
#print the table
kable(units_per_product_region, caption = "Number of Units Sold per Product per Region", 
      col.names = c("Product Name", "Region", "Total Units"))
#generate the plot
ggplot(units_per_product_region, aes(x = product_name, y = total_units, group = territory, color = territory)) +
  geom_line() +
  theme_minimal() +
  labs(title = "Units Sold per Product per Region",
       x = "Product Name", y = "Units Sold") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Analytical Query IV: Average sales per sales rep over the years.
```{r average-sales, echo=FALSE, eval=FALSE}
# Analytical Query IV: Average sales per sales rep over the years
```{r echo=FALSE}
#Query
query_avg_sales_per_rep <- "
SELECT 
    rep_name, 
    year, 
    AVG(average_amount_sold) AS avg_sales_per_year
FROM 
    rep_facts 
GROUP BY 
    rep_name, 
    year
ORDER BY 
    rep_name, 
    year;
"

avg_sales_per_rep <- dbGetQuery(con, query_avg_sales_per_rep)
avg_sales_per_rep$year <- factor(avg_sales_per_rep$year, levels = sort(unique(avg_sales_per_rep$year)))

# Generating the line plot
ggplot(avg_sales_per_rep, aes(x = year, y = avg_sales_per_year, group = rep_name, color = rep_name)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(title = "Average Sales per Sales Rep Over the Years",
       x = "Year", y = "Average Sales per Year") +
  theme(legend.position = "right") 

# Print the table
knitr::kable(avg_sales_per_rep, caption = "Average Sales per Sales Rep Over the Years")

```



```{r disconnect, echo=FALSE, eval=FALSE, message=FALSE}
# Disconnect from the database
dbDisconnect(con)
```


